Value Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking Around

### 2017-09-26 paper/paper.tex
Exciting

@@

\begin{abstract}
    At least one modern subgraph isomorphism solver use reasonably good value-ordering heuristics to
    direct search, but these heuristics are not perfect, particularly for early choices. We
    investigate two mechanisms for offsetting this weakness: discrepancy searches, and restarts. Our
    results suggest that discrepancy search is too expensive to be effective, but that restarts with
    a random value-ordering heuristic beats the best tailored heuristic. We also introduce a
    slightly-random value-ordering heuristic with similar behaviour to the tailored heuristic.  When
    combined with restarts, this new heuristic is even more effective on satisfiable instances, and
    does not substantially affect performance on unsatisfiable instances.
\end{abstract}

@@

The current single strongest subgraph isomorphism solver uses ``highest degree first'' as a
value-ordering heuristic.  This is much better than not having a value-ordering heuristic at all,
but it is far from perfect, particularly for early branching choices. There are two reasons to
suspect we could do better.  Firstly, even if degree were a perfect source of information, many
graphs do not have a large degree distribution, and so we would still need to determine \emph{which}
vertex of highest degree would be best.  Secondly, experiments with a parallel search strategy which
explicitly steals early branching choices often gives strongly super-linear speedups on satisfiable
instances.

@@

?? Quick intro to discrepancy search. Quick intro to restarts, mentioning how it's important to
change something when restarting.

@@

For the variable-ordering heuristic on \siplineref{variableordering} of \cref{algorithm:sip}, we use
smallest domain first, tiebreaking on ?? degree. Variable ordering is not the focus of this
paper---as far as possible, we will leave it intact, so as not to affect the performance of the
algorithm on unsatisfiable instances.

@@

\paragraph{Weighted} As with the random heuristic, we shuffle the vertices before we enter the
$\KwSty{foreach}$ loop. Instead of picking a vertex uniformly at random, we use a weight of
$2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$ next.\footnote{We
cheat slightly to avoid range issues with random number generation: we give the vertex with highest
degree in the entire graph weight $2^{50}$, and then give lower degree vertices smaller powers,
giving every vertex weight at least 1.  This is a bit of hack but programming is hard.} This
heuristic heavily prefers vertices of high degree, similar to the tailored heuristic, but will give
the same chance of being selected next to two vertices of equal degree. This heuristic also
introduces an element of randomness, which will be important for restarts.

@@

?? Talk about how we're not dealing with a binary search tree, so we could either make it binary, or
count one discrepancy each time the degree changes, or count every ``against'' branch as being just
one discrepancy regardless.

@@

?? We use the Luby scheme, because everyone else does. We count the number of backtracks (that is,
when we reach the end of the main $\KwSty{foreach}$ loop) to decide when to restart. Following
established wisdom, we multiply each item in the Luby sequence by a magic constant. Preliminary
experiments demonstrated that this is useful, but we failed to determine a principled way of
selecting the constant's value, so we fell back on divine revelation and set it to 666. As with
everyone else, we are not entirely clear why we are doing this: it could be to allow the solver to
spend more time deep in search, it could be to reduce the overheads from repeatedly visiting inner
nodes, or it could be to keep the nogoods small.

?? We just use two watched literals. Could use increasing nogoods etc. We are not confident enough
in our programming abilities to implement these correctly, and we are not maintaining arc
consistency anyway. We also really do not want to propagate on negative decisions, because we do not
want to iterate over every removed value.

@@

?? Dual Xeon E5-2697A v4 CPUs, 512GBytes RAM, running Ubuntu 17.04, implemented in C++ and compiled
using GCC 6.3.0. Our code is heavily based upon ??.

@@

?? The 5,725 instances.

@@

The performance of the weighted and tailored heuristics is very similar: the weighted heuristic has
slightly worse performance below a 100s timeout, and effectively equal performance above it. This
shows that we can introduce an element of randomness into the tailored value-ordering heuristic
without adversely affecting its performance.

@@

In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
the random variable-ordering heuristic comfortably beats the tailored strategy without restarts. In
other words, although having a good variable-ordering heuristic is beneficial, introducing
randomness into the search is better, if it is done alongside a mechanism to avoid heavy commitment
to any particular random choice.  However, the weighted heuristic together with restarts is better
still---that is, if we are introducing restarts, then it is better to add a small amount of
randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.

@@

    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
    versus weighted + restarts. Below, comparing instead with simple random + restarts.}
    \label{figure:scatter}
\end{figure}

Cumulative plots only show aggregate behaviour. What happens on an instance by instance basis?  In
the top of \cref{figure:scatter} we scatter plot performance, showing satisfiable and unsatisfiable
instances using dots and crosses respectively. On the $x$-axis we show the tailored heuristic,
whilst on the $y$-axis we show weighted + restarts. Points on the outer axes represent instances
which timed out with one heuristic but not the other.

@@

@@

?? Bigger CP picture

>>>>

\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
\newcommand{\citep}[1]{\cite{#1}}

@@

\begin{abstract}
    Modern subgraph isomorphism solvers use strong value-ordering heuristics to direct search, but
    these heuristics are not perfect, particularly for early choices. Our results suggest that
    discrepancy-based searches are too expensive to be effective at offsetting this weakness.
    However, restarts with a random value-ordering heuristic beats conventional search with the best
    tailored heuristic. This motivates the introduction of a slightly-random value-ordering
    heuristic with similar behaviour to the tailored heuristic: when combined with restarts, this
    new heuristic is even more effective on satisfiable instances, whilst not significantly
    decreasing performance on unsatisfiable instances.
\end{abstract}

@@

The current single strongest subgraph isomorphism solver uses ``highest degree first'' as a
value-ordering heuristic \citep{dblp:conf/cp/McCreeshP15,dblp:conf/ijcai/McCreeshPT16}.  This is
much better than not having a value-ordering heuristic at all, but it is far from perfect,
particularly for early branching choices. There are two reasons to suspect we could do better.
Firstly, even if degree were a perfect source of information, many graphs do not have a large degree
distribution, and so we would still need to determine \emph{which} vertex of highest degree would be
best.  Secondly, experiments with a parallel search strategy which explicitly steals early branching
choices often gives strongly super-linear speedups on satisfiable instances
\citep{dblp:conf/cp/McCreeshP15}.

@@

@@

For the variable-ordering heuristic on \siplineref{variableordering} of \cref{algorithm:sip}, we use
smallest domain first, tiebreaking on highest degree. Variable ordering is not the focus of this
paper---as far as possible, we will leave it intact, so as not to affect the performance of the
algorithm on unsatisfiable instances.

@@

\paragraph{Biased} As with the random heuristic, we shuffle the vertices before we enter the
$\KwSty{foreach}$ loop. Instead of picking a vertex uniformly at random, we use a weight of
$2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$ next. (We cheat
slightly to avoid range issues with random number generation: we give the vertex with highest degree
in the entire graph weight $2^{50}$, and then give lower degree vertices smaller powers, giving
every vertex weight at least 1.  This is a bit of hack but programming is hard.) This heuristic
heavily prefers vertices of high degree, similar to the tailored heuristic, but will give the same
chance of being selected next to two vertices of equal degree. This heuristic also introduces an
element of randomness, and so can be used with restarts.

@@

?? Quick intro to discrepancy search
\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/aaai/Korf96,DBLP:conf/ijcai/Walsh97,DBLP:conf/cpaior/KarouiHLN07,DBLP:journals/jea/ProsserU11}.
Discrepancy search is about value ordering.

?? Talk about how we're not dealing with a binary search tree, so we could either make it binary, or
count one discrepancy each time the degree changes, or count every ``against'' branch as being just
one discrepancy regardless. ?? Has anyone done a proper job of this?

@@

?? Quick intro to restarts, mentioning how it's important to change something when restarting.
Interestingly usually that thing is variable-ordering heuristics, not value-ordering
\citep{DBLP:conf/ijcai/LecoutreSTV07,DBLP:journals/jsat/LecoutreSTV07,DBLP:conf/cp/GayHLS15,DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17}.
?? Does anyone at all do this on value ordering explicitly? And if not, why not? There's
\citet{DBLP:conf/flairs/RazgonOP07} at least. ?? There's a connection here to confidence-based work
stealing \citep{DBLP:conf/cp/ChuSS09} too. But what? They talk about confidence in a heuristic, but
not that different values could be ranked differently.

?? We use the Luby scheme \citep{DBLP:journals/ipl/LubySZ93}, because everyone else does. We count
the number of backtracks (that is, when we reach the end of the main $\KwSty{foreach}$ loop) to
decide when to restart. Following established wisdom, we multiply each item in the Luby sequence by
a magic constant. Preliminary experiments demonstrated that this is useful, but we failed to
determine a principled way of selecting the constant's value, so we fell back on divine revelation
and set it to 666. As with everyone else, we are not entirely clear why we are doing this: it could
be to allow the solver to spend more time deep in search, it could be to reduce the overheads from
repeatedly visiting inner nodes, or it could be to keep the nogoods small.

?? We just use two watched literals \citep{DBLP:conf/dac/MoskewiczMZZM01}. Could use increasing
nogoods \citep{DBLP:conf/aaai/LeeSZ16,DBLP:conf/cp/GlorianBLLM17} etc, but we are not confident enough
in our programming abilities to implement these correctly, and we are not maintaining arc
consistency anyway. We also really do not want to propagate on negative decisions, because we do not
want to iterate over every removed value.

@@

?? Dual Xeon E5-2697A v4 CPUs, 512GBytes RAM, running Ubuntu 17.04, implemented in C++ and compiled
using GCC 6.3.0. Our code is heavily based upon \citet{dblp:conf/cp/McCreeshP15}.

@@

?? The 5,725 instances \citep{dblp:conf/lion/KotthoffMS16}.

@@

The performance of the biased and tailored heuristics is very similar: the biased heuristic has
slightly worse performance below a 100s timeout, and effectively equal performance above it. This
shows that we can introduce an element of randomness into the tailored value-ordering heuristic
without adversely affecting its performance.

@@

In \cref{figure:restarts} we show the effects of adding restarts to the algorithm. With restarts,
the random variable-ordering heuristic comfortably beats the tailored strategy without restarts. In
other words, although having a good variable-ordering heuristic is beneficial, introducing
randomness into the search is better, if it is done alongside a mechanism to avoid heavy commitment
to any particular random choice.  However, the biased heuristic together with restarts is better
still---that is, if we are introducing restarts, then it is better to add a small amount of
randomness to a tailored deterministic heuristic than it is to throw away the heuristic altogether.

@@

    \caption{Instance by instance comparison of the basic algorithm with the tailored heuristic,
    versus biased + restarts. Below, comparing instead with simple random + restarts.}
    \label{figure:scatter}
\end{figure}

Cumulative plots only show aggregate behaviour. What happens on an instance by instance basis?  In
the top of \cref{figure:scatter} we scatter plot performance, showing satisfiable and unsatisfiable
instances using dots and crosses respectively. On the $x$-axis we show the tailored heuristic,
whilst on the $y$-axis we show biased + restarts. Points on the outer axes represent instances
which timed out with one heuristic but not the other.

@@

?? maybe some engineering aspects: look at performance with restarts on tailored, and at restarts
without nogood recording

@@

?? We can keep good value-ordering heuristics, and get large improvement. Importantly, we can do
this with almost no effect on unsatisfiable instances, and rare to make satisfiable much worse.

?? Variable ordering

?? Bigger CP picture.
